Self-interacting dark matter model without dark energy in cosmology

### 2024-11-25 article/article.tex
add data

   Basilakos \& Plionis (2009) identified two functional forms for which
   the previous Boltzmann equation can be solved analytically.
   Refering to Appendix B in (), only one of these two is of interest
   because it provides a "$\propto a^{-3}$" dependence of the scale factor,
   which is \begin{equation}
      \Psi(a)=aH(a)R(a)=C_1(n+3)a^nH(a)+\kappa C_1^2a^{2m}.
   \end{equation}
   And the total energy density is
   \begin{equation}
      \rho(a)=C_1a^n+\frac{a^{-3}F(a)}{C_2-\int_1^{a}x^{-3}f(x)F(x)\mathrm{d}x},\label{eq:5}
   \end{equation}
   where $f(a)=-\kappa/[aH(a)],$ and the kernal function $F(a)$ has
   the form \begin{equation}
      F(a)=\exp\left[-2\kappa C_1\int_1^{a}\frac{x^{n-1}}{H(x)}\mathrm{d}x\right].
   \end{equation}
   The first term of Eq.(\ref{eq:5}) is the density corresponding to the
   residual matter creation that results from a possible disequilibrium
   between the particle creation and annihilation processes, while the second
   term can be viewed as the energy density of the self-IDM particles that are
   dominated by the annihilation process.

@@

   If $n=0$, the global density evolution can be transformed as
   \begin{equation}
      \rho(a)=C_1+a^{-3}\frac{e^{-2\kappa C_1(t-t_0)}}{C_2-\kappa Z(t)},
   \end{equation}
   where $Z(t)=\int_{t_0}^ta^{-3}e^{-2\kappa C_1(t'-t_0)}\mathrm{d}t'$().
   Using the usual unit-less $\Omega$-like parameterization, we obtain that
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{1,0}+\frac{\Omega_{1,0}\Omega_{2,0}a^{-3}e^{-2\kappa C_1(t-t_0)}}
      {\Omega_{1,0}+\kappa C_1\Omega_{2,0}Z(t)},
   \end{equation}
   where $\Omega_{1,0}=8\pi GC_1/3H_0^2$ and $\Omega_{2,0}=8\pi G/3H_0^2C_2$,
   which related to $\Omega_{\Lambda}$ and $\Omega_m$ in the $\Lambda$CDM model,
   respectively.
   From Eq.({\ref{eq:2}}), we can also give the mass of the DM particle
   related to the range of $\kappa C_1$(in the unit of Gyr${}^{-1}$)
   \begin{equation}
      M_x=\frac{3.325\times10^{-12}}{\kappa C_1}
      \frac{\langle\sigma u\rangle}{10^{-23}}h^2(1-\Omega_{2,0})\,\text{GeV},
   \end{equation}
   where $h\equiv H_0/[100\text{km/s/Mpc}]$.

@@

   If $\kappa=0$, the global density evolution can be written as
   \begin{equation}
      \rho(a)=\mathcal{D}a^{-3}+C_1a^{n},
   \end{equation}
   where $\mathcal{D}=C_2-C-1$. The conditions in which the current 
   model acts as a quintessence cosmology are given by $\mathcal{D}>0,
   C_1>0,$ and $w_{\text{IDM}}=-1-n/3$. This solution is mathematically
   equivalent to that of the gravitational matter creation model of().
   The Hubble flow is now given by
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{2,0}a^{-3}+\Omega_{1,0}a^{n},
   \end{equation}
   where $\Omega_{2,0}=8\pi G\mathcal{D}/3H_0^2$ and 
   $\Omega_{1,0}=8\pi GC_1/3H_0^2$, respectively.()

@@

@@

>>>>

   \cite{refId0} identified two functional forms for which
   the previous Boltzmann equation can be solved analytically.
   Refering to Appendix B in \cite{refId0}, only one of these two is of interest
   because it provides a "$\propto a^{-3}$" dependence of the scale factor,
   which is \begin{equation}
      \Psi(a)=aH(a)R(a)=C_1(n+3)a^nH(a)+\kappa C_1^2a^{2m}.
   \end{equation}
   And the total energy density is
   \begin{equation}
      \rho(a)=C_1a^n+\frac{a^{-3}F(a)}{C_2-\int_1^{a}x^{-3}f(x)F(x)\mathrm{d}x},\label{eq:5}
   \end{equation}
   where $f(a)=-\kappa/[aH(a)],$ and the kernal function $F(a)$ has
   the form \begin{equation}
      F(a)=\exp\left[-2\kappa C_1\int_1^{a}\frac{x^{n-1}}{H(x)}\mathrm{d}x\right].
   \end{equation}
   The first term of Eq.(\ref{eq:5}) is the density corresponding to the
   residual matter creation that results from a possible disequilibrium
   between the particle creation and annihilation processes, while the second
   term can be viewed as the energy density of the self-IDM particles that are
   dominated by the annihilation process.

@@

   If $n=0$, the global density evolution can be transformed as
   \begin{equation}
      \rho(a)=C_1+a^{-3}\frac{e^{-2\kappa C_1(t-t_0)}}{C_2-\kappa Z(t)},
   \end{equation}
   where $Z(t)=\int_{t_0}^ta^{-3}e^{-2\kappa C_1(t'-t_0)}\mathrm{d}t'$(\cite{refId0}).
   Using the usual unit-less $\Omega$-like parameterization, we obtain that
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{1,0}+\frac{\Omega_{1,0}\Omega_{2,0}a^{-3}e^{-2\kappa C_1(t-t_0)}}
      {\Omega_{1,0}+\kappa C_1\Omega_{2,0}Z(t)},
   \end{equation}
   where $\Omega_{1,0}=8\pi GC_1/3H_0^2$ and $\Omega_{2,0}=8\pi G/3H_0^2C_2$,
   which related to $\Omega_{\Lambda}$ and $\Omega_m$ in the $\Lambda$CDM model,
   respectively.
   From Eq.({\ref{eq:2}}), we can also give the mass of the DM particle
   related to the range of $\kappa C_1$(in the unit of Gyr${}^{-1}$)
   \begin{equation}
      M_x=\frac{3.325\times10^{-12}}{\kappa C_1}
      \frac{\langle\sigma u\rangle}{10^{-23}}h^2(1-\Omega_{2,0})\,\text{GeV},
   \end{equation}
   where $h\equiv H_0/[100\text{km/s/Mpc}]$.

@@

   If $\kappa=0$, the global density evolution can be written as
   \begin{equation}
      \rho(a)=\mathcal{D}a^{-3}+C_1a^{n},
   \end{equation}
   where $\mathcal{D}=C_2-C-1$. The conditions in which the current 
   model acts as a quintessence cosmology are given by $\mathcal{D}>0,
   C_1>0,$ and $w_{\text{IDM}}=-1-n/3$. This solution is mathematically
   equivalent to that of the gravitational matter creation model of().
   The Hubble flow is now given by
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{2,0}a^{-3}+\Omega_{1,0}a^{n},
   \end{equation}
   where $\Omega_{2,0}=8\pi G\mathcal{D}/3H_0^2$ and 
   $\Omega_{1,0}=8\pi GC_1/3H_0^2$, respectively.(\cite{refId0})

@@

   To constrain the relevant IDM models (\cite{refId0}), we use the newly revised
   observational $H(z)$ data (OHD)(\cite{Zhang_2014};\cite{PhysRevD.71.123001};
   \cite{M.Moresco_2012};\cite{Moresco_2016};\cite{10.1093/mnras/stx301};\cite{10.1093/mnrasl/slv037}),
   the Pantheon+ set of 1701 SNe Ia (), the CMB data from
   Planck 2018 and the BAO data from DESI 2024.

\subsection{The observational H(z) data}

   It is widely known that the Hubble parameter $H(z)$ depends on
   the differential age as a function of redshift $z$ in the form
   \begin{equation}
      H(z)=-\frac{1}{1+z}\frac{\mathrm{d}z}{\mathrm{d}t},
   \end{equation}
   which provides a direct measurement on $H(z)$ based on
   $\mathrm{d}z/\mathrm{d}t$.
   OHD measurements have recently been acquired mainly employing
   cosmic chronometers (CC). The CC method is used to provide 33 observational
   data points, which are taken in the redshift range [0, 1.965].
   The Table \ref{tab:1} lists the OHD dataset used in this analysis.
   In this case, $\chi^2$ can be defined as
   \begin{equation}
      \chi_{\text{OHD}}^2=\sum_i^{33}\frac{(H_{\text{th}}-H_{\text{data}})^2}{\sigma_i^2}.
   \end{equation}

   \begin{table}[htbp]
      \caption{The OHD dataset}
      \centering
      \begin{tabular}{llll}
         \hline\hline
         $z$ & $H(z)$ & 1$\sigma$ uncertainty & Reference \\
         \hline     
         0.07 & 69 & $\pm$19.6 & \cite{Zhang_2014} \\
         0.09 & 69 & $\pm$12 & \cite{PhysRevD.71.123001} \\
         0.12 & 68.6 & $\pm$26.2 & \cite{Zhang_2014} \\
         0.17 & 83 & $\pm$8 & \cite{PhysRevD.71.123001} \\
         0.179 & 75 & $\pm$4 & \cite{M.Moresco_2012} \\
         0.199 & 75 & $\pm$5 & \cite{M.Moresco_2012} \\
         0.2 & 72.9 & $\pm$29.6 & \cite{Zhang_2014} \\
         0.27 & 77 & $\pm$14 & \cite{PhysRevD.71.123001} \\
         0.28 & 88.8 & $\pm$36.6 & \cite{Zhang_2014} \\
         0.352 & 83 & $\pm$14 & \cite{M.Moresco_2012} \\
         0.3802 & 83 & $\pm$13.5 & \cite{Moresco_2016}  \\
         0.4 & 95 & $\pm$17 & \cite{PhysRevD.71.123001} \\
         0.4004 & 77 & $\pm$10.2 & \cite{Moresco_2016} \\
         0.4247 & 87.1 & $\pm$11.2 & \cite{Moresco_2016} \\
         0.4497 & 92.8 & $\pm$12.9 & \cite{Moresco_2016} \\
         0.47 & 89 & $\pm$34 & \cite{10.1093/mnras/stx301} \\
         0.4783 & 80.9 & $\pm$9 & \cite{Moresco_2016} \\
         0.48 & 97 & $\pm$62 & \cite{Daniel.Stern_2010} \\
         0.593 & 104 & $\pm$13 & \cite{M.Moresco_2012} \\
         0.68 & 92 & $\pm$8 & \cite{M.Moresco_2012} \\
         0.75 & 98.8 & $\pm$33.6 & \cite{Borghi_2022} \\
         0.781 & 105 & $\pm$12 & \cite{M.Moresco_2012} \\
         0.8 & 113.1 & 15.1 & \cite{Jiao_2023} \\
         0.875 & 125 & $\pm$17 & \cite{M.Moresco_2012} \\
         0.88 & 90 & $\pm$40 & \cite{Daniel.Stern_2010} \\
         0.9 & 117 & $\pm$23 & \cite{PhysRevD.71.123001} \\
         1.037 & 154 & $\pm$20 & \cite{M.Moresco_2012} \\
         1.3 & 168 & $\pm$17 & \cite{PhysRevD.71.123001} \\
         1.363 & 160 & $\pm$33.6 & \cite{10.1093/mnrasl/slv037} \\
         1.43 & 177 & $\pm$18 & \cite{PhysRevD.71.123001} \\
         1.53 & 140 & $\pm$14 & \cite{PhysRevD.71.123001} \\
         1.75 & 202 & $\pm$40 & \cite{PhysRevD.71.123001} \\
         1.965 & 186.5 & $\pm$50.4 & \cite{10.1093/mnrasl/slv037} \\
         \hline    
      \end{tabular}
      \label{tab:1}
   \end{table}

\subsection{The observational SNe Ia data}

   SNe Ia have long been used as "standard candles". It is

@@

\bibliographystyle{aa}
\bibliography{reference}
